\documentclass[10pt,a4paper]{amsart}

\usepackage{lmodern}
\linespread{1.1}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{fancyref}
\usepackage[colorlinks=true]{hyperref}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{tikz-cd}

\newtheorem{theo}{Theorem}[section]
\newtheorem{prop}[theo]{Proposition}
\newtheorem{lemm}[theo]{Lemma}
\newtheorem{coro}[theo]{Corollary}

\theoremstyle{definition}
\newtheorem{defi}[theo]{Definition}
\newtheorem{exam}[theo]{Example}
\newtheorem*{claim}{Claim}

\newcommand{\kk}[1]{\mathbb{#1}}
\newcommand{\cc}[1]{\mathcal{#1}}

\def\qedhere{}

\def\^#1{^{[#1]}}
\def\qandq{\quad\text{and}\quad}
\def\ov#1{\overline{#1}}

\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\im}{Im}
\DeclareMathOperator{\Vol}{Vol}
\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\tr}{tr}

\def\snd{\sigma}

\newcommand{\ext}[1]{\bigwedge{}^{\!\!#1}\,}

\newtheorem{question}{Question}

\author{Gunnar Þór Magnússon}
\date{\today}
\title[Degenerate Hermitian geometry]{Degenerate Hermitian geometry\\
and curvatures of holomorphic fibrations}

\hypersetup{
 pdfauthor={Gunnar Þór Magnússon},
 pdftitle={Degenerate Hermitian geometry},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.1 (Org mode 9.3)},
 pdflang={English}}


\begin{document}



\begin{abstract}
We calculate curvature tensors of metrics on the total spaces of holomorphic fibrations. Our main tool is a theory of Chern connections and curvature forms for possibly degenerate Hermitian forms on holomorphic vector bundles. We prove a version of the Codazzi--Griffiths equations for curvatures of sub- and quotient bundles in that setting and apply them to the study of honest Hermitian metrics on fibrations. Applications include a calculation of the curvature tensor of a metric on Grassmannian bundles.
\end{abstract}


\maketitle






\section*{Introduction}

Let $\pi : X \to S$ be a family of complex manifolds, and consider the associated short exact sequence
\[
\begin{tikzcd}
0 \ar[r] & T_{X/S} \ar[r] & T_X \ar[r] & \pi^*T_S \ar[r] & 0.
\end{tikzcd}
\]
We often find ourselves in the situation of having a metric $h_S$ on the base $S$ and a Hermitian form $h_{X/S}$ on the $X$ that is positive-definite on the relative tangent bundle $T_{X/S}$, and then construct a metric on $X$ as the sum $h_X := h_{X/S} + \pi^*h_S$. This happens when we consider a projective or Grassmannian bundle associated to a vector bundle~\cite{alvarez2016positive,alvarez2018projectivized,yang2019hirzebruch}; the blow-up of a point like in the original proof of the Kodaira embedding theorem~\cite{kodaira-embedding}; metrics on holomorphic fibrations~\cite{calabi-fibres-holomorphes}; or metrics on versal families of deformations~\cite{magnusson2012natural}.

In these situations, the metric on the base has a curvature tensor, and the relative form often has something that resembles such a tensor. As neither form is positive-definite on $X$, we can't use our usual tools to calculate a curvature tensor for $h_X$ from this information. Instead, we are often reduced to committing difficult calculations in local coordinates; see for example \cite{chaturvedi2019hermitian}.

If we had the equivalent of Chern connections and curvature forms for arbitrary Hermitian forms, we could try to prove a version of the Codazzi--Griffiths equations~\cite{griffiths1965hermitian} that relate the curvature form of sub- and quotient bundles to that of the ambient bundle. In the positive-definite case, those equations can be used to calculate the curvature tensor of the sum $h_1 + h_2$ of two metrics on a given bundle $E$ by considering the short exact sequence
\[
\begin{tikzcd}
0 \ar[r] & E \ar[r] & E\oplus E \ar[r] &  E \ar[r] & 0,
\end{tikzcd}
\]
where the metrics on the bundles are (in order) $h_1 + h_2$, $h_1 \oplus h_2$, and the induced metric on the ``quotient''.

In this note, we fill in the details of this sketch. In section~\ref{sec:degenerate-chern-connections} we show that a holomorphic vector bundle $E \to X$ with a smooth Hermitian form $b$ admits a Chern connection. Such a connection is not unique when the form is degenerate, but we can always get rid of the ambiguity between different choices of connections by applying the form $b$, as the difference of two such connections must take values in its kernel. This yields a unique curvature \emph{tensor} for such a form. We then prove the Codazzi--Griffiths equations hold in this setting, and turn to applications.

The first is a dotting of i's and crossing of t's from works of \'Alvarez, Heier and Zheng~\cite{alvarez2018projectivized,alvarez2016positive} on the curvature of metrics on projective bundles (see also Yang and Zheng~\cite{yang2019hirzebruch}). They prove that such bundles admit metrics of positive holomorphic sectional curvature, and finish by saying the same arguments can prove that so do Grassmannian bundles. We treat the case of Grassmannian bundles in detail, and as a corollary conclude that bundles of flag manifolds admit metrics of positive holomorphic sectional curvature.

The second application is the paying of a debt of an unfulfilled expectation of the author's Ph.D.~advisor. Consider a family $\pi : X \to S$ of compact K\"ahler manifolds over a smooth base. We can complexify the K\"ahler cone of each manifold and obtain a holomorphic fibration $p : C \to S$ of such cones. Hodge theory gives a natural smooth semipositive Hermitian form on $C$ that is positive-definite on the relative tangent bundle $T_{C/S}$. Given a metric on the base, we can thus construct a metric on $C$, and we use the tools developed here to calculate its curvature tensor. We pay special attention to the case of compact K\"ahler manifolds with zero first Chern class, whose base of versal deformations is smooth and admits a Weil--Petersson metric.

To do this, we require a degenerate analogue of many statements about adjoints, orthogonal subspaces and induced inner products that are classical in the positive-definite case. We do not know a reference for these results, which would make for fun exercises in a linear algebra course, and thus provide statements and proofs. We do this in section~\ref{sec:degenerate-linear-algebra} as we need to refer to those results for our work on connections and curvature.

A seeming novelty of our approach is that we prove all our results using the connection formulation of complex differential geometry (versus the local coordinate or moving frame formulations). While this connection approach is common in modern Riemannian geometry, complex geometers seem to prefer the older, index-heavy, ways. We hope this note also serves as an advertisement for the advantages of the connection approach.



We finish this introduction by noting that we think the technical results developed here on general Hermitian forms are only interesting when applied to the study of honest Hermitian metrics, and do not have merit on their own.
Some examples may explain this opinion:


\begin{exam}
More degenerate the form is, the less information the connection gives:
Let $E \to X$ be a holomorphic vector bundle, let $b = 0$, and let $D$ be any connection whose $(0,1)$-part is $\bar\partial$. Then $D$ is a Chern connection of $b$.
\end{exam}


\begin{exam}
Let $X$ be a compact complex manifold of dimension $n$
that admits a nowhere-zero holomorphic one-form $\alpha$.
It  gives us a smooth Hermitian form $b = i \alpha \wedge \ov \alpha$ whose kernel forms a holomorphic subbundle of rank $n-1$ and whose quotient is the trivial line bundle.
Any of our Chern connections for this form will yield a curvature form whose top exterior power vanishes; so if we were to develop Chern--Weil theory in this setting we should expect to conclude the top Chern class of the manifold is zero.
However, a much easier way to conclude the same is to use the additivity property of total Chern classes on the short exact sequence we began with.
\end{exam}


\begin{exam}
Even when a form is non-degenerate, having a Chern connection and curvature form just doesn't let us say very much:
Let $\pi : X \to S$ be a family of compact complex K\"ahler manifolds of dimension $2n$ over a smooth base $S$.
Let $E \to S$ be the holomorphic vector bundle whose sheaf of sections is $\mathcal R^{n}\pi_{*}\mathbb C \otimes \mathcal O_{S}$ and whose fiber over a point $s$ is $E_{s} = H^{n}(X_{s}, \mathbb C)$.
This vector bundle admits a non-degenerate Hermitian form given by the cap product, and the flat Gauss--Manin connection is the Chern connection of this Hermitian form.
\end{exam}

Also note that in the non-degenerate case the most useful results are found when the form is positive-definite. For example the Laplacian of a non-degenerate form of mixed signature will not be strongly elliptic, so the Hodge isomorphism theorem fails, and with it go most of the tools and techniques developed over the last hundred years or so.

We'll finish with one final example we don't have any applications in mind for, but find amusing.

\begin{exam}
Let $(E,h) \to X$ be a holomorphic vector bundle with a Hermitian metric $h$. The curvature tensor of the metric,
\[
  R(\alpha, \ov\beta, s, \ov t)
  = h(\tfrac i2 \Theta_{\alpha, \ov\beta}s, \ov t),
\]
defines a Hermitian form on $E \otimes T_{X}$ that is usually not positive-definite. Good luck saying anything about its curvature tensor.
\end{exam}



\section{Linear algebra}
\label{sec:degenerate-linear-algebra}



All vector spaces are finite-dimensional and over the complex numbers. Let $V$ be a vector space and let $b$ be a Hermitian form on $V$. We want to study the geometry of the pair $(V,b)$.\footnote{Everything we say also applies to symmetric bilinear forms on finite-dimensional vector spaces over arbitrary fields. The reader who is interested in those can mentally erase all the conjugations of vectors or spaces.}
Our objective is to establish a version of the Codazzi--Griffiths equations for sub- and quotient bundles of a holomorphic vector bundle, so the main questions we want to answer here are:
\begin{itemize}
\item
When does an adjoint of a linear morphism exist?

\item
Does a Hermitian form on $V$ induce Hermitian forms on all spaces derived from its tensor algebra?

\item
Given a short exact sequence $0 \to S \to V \to Q \to 0$ and a Hermitian form on $V$, do we get Hermitian forms on $S$ and $Q$?
\end{itemize}


\begin{defi}
A \emph{Hermitian space} is a finite-dimensional vector space $V$ with a Hermitian form $b$.
A \emph{morphism} of Hermitian spaces is a linear morphism $f : V \to W$ such that
\[
b_W(fx, \ov{fy}) = b_V(x, \ov y)
\]
for all $x, y \in V$.
\end{defi}

The category this forms is the one that has Hermitian spaces for objects and the above morphisms between them. Given two objects in this category, there is not necessarily any morphism between them; if $V$ has a non-degenerate form while $W$ has the zero form, there is no morphism $V \to W$.

A Hermitian form can be seen as a linear morphism $b : V \to \ov V^*$. As such, it has a kernel. We clearly have
\[
\Ker b = \{ x \in V \mid b(x, \ov y) = 0 \text{ for all $y \in V$}\}.
\]


The results we want to prove are all known for spaces with a non-degenerate form. Our main technical tool will be to reduce to that case by quotienting out the kernel of the form we have.

\begin{prop}
Let $V$ be a vector space. For every Hermitian form $b$ on $V$, there exists a unique non-degenerate Hermitian form $\hat b$ on $V / \Ker V$ such that $q : V \to V/\Ker b$ is a Hermitian morphism.
\end{prop}


\begin{proof}
We define
\[
\hat b(qx, \ov{qy})
= b(x, \ov y).
\]
This is well-defined, as any other element in $qx$ differs from the one we chose by an element of $\Ker q = \Ker b$. The form $\hat b$ is also non-degenerate, as if $qx \in \Ker \hat b$ then $x \in \Ker b$, so $qx = 0$. The form is also defined in such a way that $q$ is a morphism of Hermitian spaces.

If $b'$ is another non-degenerate Hermitian form on $V / \Ker b$ that makes $q$ into a Hermitian morphism, then
\[
b'(qx, \ov{qy})
= b(x, \ov{y})
= b(qx, \ov{qy})
\]
for all $qy \in V / \Ker b$. Thus $b(qx) = b'(qx)$ as $(0,1)$-forms for all $qx \in V / \Ker b$, so $b = b'$.
\end{proof}


If this operation should be called anything, it is clearly a \emph{purge} of the original space, as it removes all the degeneracy from the Hermitian form.



\subsection*{Existence of adjoints}
\label{sec:existence-adjoints}




\begin{defi}
Let $(V, b_V)$ and $(W, b_W)$ be Hermitian spaces, and let $f : V \to W$ be a linear morphism. An \emph{adjoint} of $f$ is a linear morphism $f^\dagger : W \to V$ such that
\[
b_V(f^\dagger x, \ov y)
= b_W(x, \ov{f y})
\]
for all $x \in W$ and $y \in V$.
\end{defi}


If $b_V$ is non-degenerate, every morphism $f : V \to W$ admits a unique adjoint. It is defined by
\[
f^\dagger = b_V^{-1} \circ \ov f^* \circ b_W,
\]
where $\ov f^*$ is the conjugate dual morphism $f^* : \ov W^* \to \ov V^*$ that $f$ induces. In general adjoints may not exist.


\begin{theo}
Let $f : V \to W$ be a linear morphism. The morphism $f$ admits an adjoint if and only if $f(\Ker b_V) \subset \Ker b_W$. If the set of adjoints of $f$ is not empty, it is a $\Hom(W, \Ker b_V)$ torsor.
\end{theo}

\begin{proof}
If $f$ admits an adjoint $f^\dagger$, then
\[
b_V(f^\dagger x, \ov y)
= b_W(x, \ov{fy})
\]
for all $x \in W$ and $y \in V$. Taking $y \in \Ker b$, we see that $b_W(x, \ov{fy}) = 0$ for all $x \in W$, so $f(y) \in \Ker b_W$.

Suppose then that $f(\Ker b_V) \subset \Ker b_W$. Then $f$ defines a linear morphism $\hat f : V / \Ker b_V \to W / \Ker b_W$. As the induced forms on those spaces are non-degenerate, we get a unique adjoint $\hat f^\dagger : W / \Ker b_W \to V / \Ker b_V$. It induces a morphism $W \to V / \Ker b_V$. Picking a lift $V / \Ker b_V \to V$ we then get an adjoint $f^\dagger$ of $f$.

If $g$ and $h$ are adjoints of $f^\dagger$, then
\[
b_V((g - h)x, \ov y)
= b_V(gx, \ov y) - b_V(hx, \ov y)
= b_W(x, \ov{fy}) - b_W(x, \ov{fy})
= 0
\]
for all $x \in W$ and $y \in V$, so the image of $g - h$ is contained in $\Ker b_V$.
\end{proof}


\begin{coro}
The subset of $\Hom(V,W)$ of morphisms that have adjoints is a linear subspace
of codimension $\dim \Ker b_V (\dim W - \dim \Ker b_W)$.
\end{coro}

\begin{proof}
The subset is not empty because the zero morphism always has an adjoint,
and it is clearly a subspace.

Let's pick splittings $V = \Ker b_V \oplus V / \Ker b_V$ and $W = \Ker b_W
\oplus W / \Ker b_W$. Then we get a splitting
$$
\displaylines{
\Hom(V, W)
= \Hom(\Ker b_V, \Ker b_W)
\oplus \Hom(\Ker b_V, W / \Ker b_W)
\hfill\cr\hfill{}
\oplus \Hom(V / \Ker b_v, \Ker b_W)
\oplus \Hom(V / \Ker b_v, W / \Ker b_W).
}
$$
A morphism $f : V \to W$ satisfies $f(\Ker b_V) \subset f(\Ker b_W)$ if and only
if its projection onto $\Hom(\Ker b_V, W / \Ker b_W)$ in this splitting is zero.
\end{proof}



\begin{coro}
If $f : V \to W$ has an adjoint $f^\dagger$, then $f^\dagger$ also has an adjoint.
\end{coro}

\begin{proof}
If $x \in \Ker W$, then
\[
b_V(f^\dagger x, \ov y)
= b_W(x, \ov{fy})
= 0
\]
for all $y \in V$, so $f^\dagger x \in \Ker b_V$.
\end{proof}

The difference between a morphism $f$ and an adjoint of its adjoint is an element of $\Hom(V, \Ker b_V)$, so a double adjoint is unique and equal to $f$ if and only if $b_V$ is non-degenerate.


\begin{prop}
Let $j : S \to V$ be a linear morphism, let $b_V$ be a Hermitian form on $V$, and set $b_S = j^*b_V$. Then $j$ admits an adjoint.
\end{prop}

\begin{proof}
We have
\[
\Ker b_S = j^{-1}(\Ker b_V) + \Ker j
\]
so $j(\Ker b_S) = j(j^{-1}(\Ker b_V)) \subset \Ker b_V$.
\end{proof}






\subsection*{Orthogonal complements}


The objective of this section is to show how to obtain a Hermitian form on a quotient space given one on the ambient space. It is well-known how to do this when the forms are non-degenerate (for example, take duals to embed the quotient dual in the ambient dual, restrict the dual metric, and invert) but less clear how this should work in the general case.


\begin{defi}
Let $(V,b)$ be a Hermitian space, and let $S \subset V$ be a subspace. The \emph{orthogonal complement} of $S$ is
\[
S^\perp = \{ v \in V \mid b(v, \ov w) = 0 \text{ for all $w \in S$}\}.
\]
\end{defi}

The orthogonal complement of a subspace is clearly again a subspace.

\begin{prop}
\[
\Ker b \subset S^\perp.
\]
\end{prop}

\begin{proof}
Let $v \in \Ker b$. Then $b(v, \ov w) = 0$ for all $w \in S$.
\end{proof}


As before we have the quotient space $q : V \to V / \Ker b$, and the induced non-degenerate form $\hat b$ on $V / \Ker b$.

\begin{prop}
Let $S \subset V$. Then
\[
q(S^\perp) = q(S)^\perp,
\]
where the orthogonal complement on the right-hand side is with respect to $\hat b$.
\end{prop}

\begin{proof}
Let $v \in S^\perp$, so $b(v, \ov w) = 0$ for all $w \in S$. Then $\hat b(qv, \ov{qw}) = 0$ for all $w \in S$, which means that $\hat b(qv, \ov w) = 0$ for all $w \in q(S)$, so $qv \in q(S)^\perp$.

Suppose then that $v \in q(S)^\perp$, so $\hat b(v, \ov w) = 0$ for all $w \in q(S)$. If $x \in V$ is such that $qx = v$, then this means that $b(x, \ov w) = 0$ for all $w \in S$. Then $x \in S^\perp$, so $qx \in q(S^\perp)$.
\end{proof}



\begin{prop}
\[
S + S^\perp = V.
\]
\end{prop}

\begin{proof}
We have
\[
q(S + S^\perp)
= q(S) + q(S^\perp)
= q(S) + q(S)^\perp
= V / \Ker b
\]
as $\hat q$ is non-degenerate. As $q$ is surjective, this implies the result.
\end{proof}




\begin{prop}
\[
S \cap S^\perp = S \cap \Ker b.
\]
\end{prop}

\begin{proof}
We have
\[
q(S \cap S^\perp)
= q(S) \cap q(S^\perp)
= q(S) \cap q(S)^\perp
= 0
\]
as $\hat q$ is non-degenerate,
so $S \cap S^\perp \subset \Ker q = \Ker b$. Conversely, $\Ker b \subset S^\perp$, so $S \cap \Ker b \subset S \cap S^\perp$.
\end{proof}


If $f : S \to V$ is a linear morphism and $b_V$ is a Hermitian form on $V$, then
we get an induced Hermitian form $b_S := f^*b_V$ on $S$. We can ask whether the
same happens when we have a morphism $f : V \to Q$ and a form $b_V$. Clearly we
can only expect something on the image of $f$, so we should take it to be
surjective. When the form on $b_V$ is positive-definite, we can define a form on
$Q$ by using the injection $0 \to Q^* \to V^*$ and the dual Hermitian forms. In
general, we can still get a Hermitian form on the quotient space by using
orthogonal complements.


\begin{theo}
Let $q : V \to Q$ be a surjective morphism. A Hermitian form $b_V$ on $V$ induces a Hermitian form $b_Q$ on $Q$.
\end{theo}

\begin{proof}
Let $S = \Ker q$, and consider $S^\perp \subset V$. We have a short exact sequence
\[
\begin{tikzcd}
0 \ar[r] &
S \cap \Ker b_V \ar[r] &
S^\perp \ar[r,"q"] &
Q \ar[r] &
0
\end{tikzcd}
\]
because $S^\perp \cap \Ker q = S^\perp \cap S = S \cap \Ker b_V$, and $q$ restricted to $S^\perp$ is still surjective. We define
\[
b_Q(qx, \ov{qy})
= b_V(x, \ov y)
\]
for $x, y \in S^\perp$. This is well-defined by the above, as any other element $x'$ that maps to $qx$ satisfies $x' - x \in S \cap \Ker b_V \subset \Ker b_V$.
\end{proof}



\begin{prop}
Let $q : V \to Q$ be surjective, and let $b_V$ and $b_Q$ be as above. Then $q$ admits an adjoint $q^\dagger : Q \to V$.
\end{prop}

\begin{proof}
We need to show that $q(\Ker b_V) \subset \Ker b_Q$. If $x, y$ are elements of $V$, we have by definition
\[
b_Q(qx, \ov{qy}) = b_V(x, \ov y).
\]
If $x \in \Ker b_V$, this implies that $b_Q(qx, \ov{qy}) = b_V(x, \ov y) = 0$ for all $y \in V$, so $qx \in \Ker b_Q$.
\end{proof}





\subsection*{Induced forms on tensor algebras}
\label{sec:induced-forms-tensor}


If $V$ is a vector space with a Hermitian form $b_V$, we get induced Hermitian forms on any vector space derived from the tensor algebra of $V$ if we restrict ourselves to spaces that only involve tensor powers of $V$ and not $V^*$. Handling $V^*$, or more generally $\Hom(V,W)$, is more delicate when the forms are possibly degenerate.



If $V$ and $W$ are vector spaces with Hermitian forms $b_V$ and $b_W$, we'd like to use those forms to define a Hermitian form on $\Hom(V,W)$. When the form $b_V$ is non-degenerate, every linear morphism $f : V \to W$ admits a unique adjoint $f^\dagger : W \to V$, and the Fr\"obenius form on $\Hom(V,W)$ is
\[
b(f, \ov g) = \tr(g^\dagger f).
\]
In general, not every morphism $f : V \to W$ admits an adjoint, and the ones that do may not admit unique adjoints. We could try to define a Hermitian form on the subspace of morphisms that admit adjoints by $\tr(g^\dagger f)$, but as adjoints may not be unique this is not well-defined.

The way around this difficulty is again to pass to the quotient spaces. A morphism $f : V \to W$ admits an adjoint if and only if $f(\Ker b_V) \subset \Ker b_W$. In the latter case, there exists a unique morphism $\hat f : V / \Ker b_V \to W / \Ker b_W$ such that the diagram
\[
\begin{tikzcd}
V \ar[d,"\pi_V"] \ar[r,"f"] & W \ar[d,"\pi_W"]
\\
V / \Ker b_V \ar[r,"\hat f"] & W / \Ker b_W
\end{tikzcd}
\]
commutes; it is defined by $\hat f([v]) = \pi_W(f(v))$.


\begin{defi}
Let $(V,b_V)$ and $(W,b_W)$ be Hermitian spaces. We define
\[
b_{\Hom(V,W)}(f, \ov g) := \tr((\hat g)^\dagger \hat f)
\]
on the subspace of $\Hom(V,W)$ of morphisms that admit adjoints.
\end{defi}

If we denote by $\Hom_b(V,W) \subset \Hom(V,W)$ the subspace of morphisms that admit adjoints, the definition says that we have a surjective morphism
\[
\Hom_b(V,W) \to \Hom(V/\Ker b_V, W/\Ker b_W)
\]
and that we define a Hermitian form upstairs by pullback of the one downstairs, which exists because the Hermitian forms there are non-degenerate.

The special case $V^* = \Hom(V, \kk C)$, where $\kk C$ is equipped with the
standard inner product, is interesting.
Here an element $\lambda \in V^*$ admits an adjoint if and only if $\lambda(\Ker b_V)
= 0$, that is, if $\Ker b_V \subset \Ker \lambda$.
Note however that the dimension of the space of dual vectors that admit adjoints
is $\dim V - \dim \Ker b_V$.
Meanwhile, the dimension of the space of morphisms $V \to W$ that admit adjoints
is $\dim V \dim W - \dim \Ker b_V (\dim W - \dim \Ker b_W)$.
We then have
$$
\dim \Hom_b(V, W)
- \dim V^*_b \dim W
= \dim \Ker b_v \dim \Ker b_W,
$$
which is generally positive, so the canonical map $V_b^* \otimes W \to \Hom_b(V,
W)$ is generally only injective and not bijective like in the non-degenerate
case.
The problem is that the image of any element on the left-hand side will be a
morphism $f$ that satisfies $f(\Ker b_V) = 0$, while there are more elements on
the right-hand side.


\section{Chern connections}
\label{sec:degenerate-chern-connections}



Let $E \to X$ be a holomorphic vector bundle with a smooth Hermitian form $b$. We say that a connection $D$ on $E$ is \emph{compatible with $b$} if
\[
d b(s, \ov t) = b(Ds, \ov t) + b(s, \ov{Dt})
\]
for all sections $s, t$ of $E$. By composing a connection with the projection onto $(1,0)$- or $(0,1)$-forms, we get the $(1,0)$- and $(0,1)$-parts of the connection.


\begin{prop}
Let $E \to X$ be a holomorphic vector bundle and let $b$ be a smooth Hermitian form on $E$. Then there exists a connection $D$ on $E$ that's compatible with $b$ and has $D^{0,1} = \bar\partial$.
\end{prop}


\begin{proof}
The plan is to construct a connection locally on trivializing neighborhoods of $E$ and to glue the pieces together with a partition of unity.

Let $s$ be a smooth section of $E$ over a neighborhood $U$. We want to solve the equation
\begin{equation}
\label{eq:conn}
\partial_\alpha b(s, \ov t)
= b(A(\alpha), \ov t),
\end{equation}
where $\alpha \in \cc C^\infty(U, T_X)$, $t \in \cc C^\infty(U, E)$ and $A \in \cc A^{1,0}(U, E)$. Consider the short exact sequence
\[
\begin{tikzcd}
0 \ar[r] &
\Ker b \ar[r] &
E \ar[r,"b"] &
\im b \ar[r] &
0,
\end{tikzcd}
\]
and tensor it by the vector bundle $\Omega^1_X$ to obtain
\[
\begin{tikzcd}
0 \ar[r] &
\Omega^1_X \otimes \Ker b \ar[r] &
\Omega^1_X \otimes E \ar[r,"\id \otimes b"] &
\Omega^1_X \otimes \im b \ar[r] &
0.
\end{tikzcd}
\]
We claim that $\alpha \otimes t \mapsto \partial_\alpha b(s, \ov t)$ is a smooth section of $\Omega^1_X \otimes \im b$. By the above, this is true if it is zero on the sheaf generated by sections of the form $\alpha \otimes t$, where $t \in \Ker b$. But for such a section we have $b(s, \ov t) = 0$ on $U$, so $\partial_\alpha b(s, \ov t) = 0$. By taking a small enough neighborhood so the sheaf morphism above is surjective, or noting that the sheaves are ones of smooth modules which are fine, we obtain a section $A_s \in \cc C^\infty(U, \Omega_X^1 \otimes E)$ such that \eqref{eq:conn} holds for any $\alpha$ and $t$.

Suppose now that there exists a holomorphic frame $(e_1, \ldots, e_r)$ of $E$ over $U$. Such a frame exists over a small enough neighborhood around any point. Let $A_1, \ldots, A_r$ be the sections we just found for $e_1, \ldots, e_r$. Any section $s \in \cc C^\infty(U, E)$ can be written as $s = \sum_j f_j e_j$ for smooth functions $f_j$ on $U$. We define
\[
D s := \sum_{j=1}^r df_j \otimes e_j + f_j A_j
\]
on $U$. We claim that $D$ is a connection on $E$ over $U$ that is compatible with $b$ and whose $(0,1)$-part is $\bar\partial$.

This object $D$ is a $1$-form with values in $E$ by construction of the $A_j$. If $s$ is a section of $E$ and $f$ a smooth function, we write $s = \sum_j f_j e_j$. Then
\begin{align*}
D(fs)
&= \sum d(f f_j) \otimes e_j + f f_j A_j
\\
&= \sum df \otimes f_j e_j + f df \otimes e_j + f f_j A_j
= df \otimes s + f Ds,
\end{align*}
so $D$ is a connection on $E$. The forms $A_j$ are $(1,0)$-forms by construction, so the $(0,1)$-part of $Ds$ is $\sum_j \bar\partial f_j e_j = \bar\partial s$.

Finally, let $f$ and $g$ be smooth functions. We have
\begin{align*}
d b(fe_j, \ov{ge_k})
&= d\bigl(f \ov g b(e_j, \ov e_k) \bigr)
\\
&= df \cdot \ov g b(e_j, \ov e_k)
+ f dg \cdot b(e_j, \ov e_k)
+ \partial b(e_j, \ov e_k)
+ \ov{\partial b(e_k, \ov e_j)}
\\
&= b(df \otimes e_j, \ov{ge_k})
+ b(f e_j, \ov{dg \otimes e_k})
+ b(A_j, \ov e_k)
+ b(e_j, \ov{A_k})
\\
&= b(D(f e_j), \ov{ge_k}) + b(f e_j, \ov{D(g e_k)}).
\end{align*}
For smooth sections $s = \sum f_j e_j$ and $t = \sum g_k e_k$ we deduce from the above and linearity that $D$ is compatible with $b$ over the neighborhood $U$.

We now cover $X$ by neighborhoods $U_\mu$ as above and construct connections $D_\mu$ on each $U_\mu$. We then take a partition of unity $(\phi_\mu)$ relative to this covering, and define a connection $D$ on $E$ by setting
\[
D s = \sum_\mu D_\mu(\phi_\mu s).
\]
This defines a connection on $E$ that's compatible with $b$ and whose $(0,1)$-part is $\bar\partial$.
\end{proof}



We will call a connection $D$ that's compatible with $b$ and has $D^{0,1} = \bar\partial$ a \emph{Chern connection} of $b$. If $b$ has a nontrivial kernel, such a connection is not unique.


\begin{prop}
Let $D_1$ and $D_2$ be Chern connections of $b$. Then $D_1 - D_2$ is an element of $\cc A^{1,0}(\Hom(E, \Ker b))$.
\end{prop}

\begin{proof}
It is well-known that the difference of two connections is an element of $\cc A^1(\End E)$. Both connections here have the same $(0,1)$-part, so their difference is a $(1,0)$-form. If $s$ and $t$ are holomorphic sections, we have
\[
b((D'_1 - D'_2)s, \ov t)
= \partial b(s, \ov t) - \partial b(s, \ov t) = 0,
\]
so $D_1 - D_2 = D_1' - D_2'$ takes values in $\Ker b$.
\end{proof}


\begin{prop}
If $s$ is a section of $\Ker b$ and $D$ is a Chern connection of $E$, then $Ds$ takes values in $\Ker b$.
\end{prop}

\begin{proof}
If $s$ is a section of $\Ker b$, then
\[
0
= \partial b(s, \ov t)
= b(Ds, \ov t)
\]
for all holomorphic sections $t$ of $E$, so $D s \in \cc A^1(\Ker b)$.
\end{proof}


\begin{theo}
Let $(E, b) \to X$ be a holomorphic vector bundle with a Hermitian form, and let $D$ be a Chern connection of $E$. The curvature tensor
\[
R(\alpha,\ov\beta,s, \ov t)
= b\bigl(\tfrac i2 D^2_{\alpha\ov\beta}s, \ov t\bigr)
\]
is independent of the choice of $D$.
\end{theo}

\begin{proof}
Let $D_1$ and $D_2$ be Chern connections of $b$. Then $D_1 = D_2 + A$, where $A \in \Hom(E, \Ker b)$, and
\[
D_1^2
= (D_2 + A)^2
= D_2^2 + A D_2 + D_2 A + A^2.
\]
The difference $D_1^2 - D_2^2$ thus takes values in $\Ker b$.
\end{proof}




\begin{prop}
Let $(E,b_E) \to X$ and $(F,b_F) \to X$ be holomorphic vector bundles with Hermitian forms and let $D_E$ and $D_F$ be Chern connections of $E$ and $F$. Let $f : E \to F$ be a smooth bundle morphism that admits an adjoint $f^\dagger$. Then $D_{\Hom(E,F)}f$ admits an adjoint and $(D_{\Hom(E,F)}f)^\dagger$ and $D_{\Hom(F,E)}f^\dagger$ are both its adjoints.
\end{prop}


\begin{proof}
The morphism $f$ admits an adjoint if and only if $f(\Ker b_E) \subset \Ker b_F$. Let $s$ be a section of $\Ker b_E$, so $f(s)$ is a section of $\Ker b_F$. Then
\begin{align*}
0 = d b_F(f(s), \ov t)
&= b_F(D_F(f(s)), \ov t) + b_F(f(s), \ov{Dt})
\\
&= b_F(D_F(f(s)), \ov t)
\\
&= b_F(D_{\Hom(E,F)}f(s), \ov t) + b_F(f(D_Es), \ov t)
\\
&= b_F(D_{\Hom(E,F)}f(s), \ov t)
\end{align*}
for all sections $t$, as $D_Es$ is a one-form with values in $\Ker b_E$. Then $D_{\Hom(E,F)} f(s) \in \cc A^1(\Ker b_F)$ for all sections $s$ of $\Ker b_E$, so $D_{\Hom(E,F)}f$ admits an adjoint.

We have $b_F(f(s), \ov{t}) = b_E(s, \ov{f^\dagger t})$ for all sections $s,t$. Taking the exterior derivative of both sides we get
\[
b_F(D_F(f(s)), \ov t) + b_F(f(s), \ov{D_F t})
= b_E(D_E s, \ov {f^\dagger t}) + b_E(s, \ov{D_E(f^\dagger t)}).
\]
The left-hand side equals
\[
b_F(D_{\Hom(E,F)}f(s), \ov t)
+ b_F(f(D_Es), \ov t)
+ b_E(s, \ov{f^\dagger(D_F t)})
\]
while the right-hand side equals
\[
b_E(f(D_E s), \ov t)
+ b_E(s, \ov{D_{\Hom(F,E)}f^\dagger(t)})
+ b_E(s, \ov{f^\dagger(D_F t)})
\]
so after cancelling terms we have
\[
b_E(s, \ov{D_{\Hom(F,E)}f^\dagger(t)})
= b_F(D_{\Hom(E,F)}f(s), \ov t)
= b_E(s, \ov{(D_{\Hom(E,F)}f)^\dagger(t)})
\]
for all sections $s,t$.
\end{proof}

When $b_E$ is non-degenerate morphisms from $E$ to $F$ admit a unique adjoint, so in that case $D_{\Hom(F,E)}f^\dagger = (D_{\Hom(E,F)}f)^\dagger$. Note that taking the adjoints of forms with values in a bundle conjugates the form part.



\begin{prop}
Let $f : X \to Y$ be a holomorphic morphism and $E \to Y$ be a holomorphic vector bundle.
Let $b$ be a Hermitian form on $E$ with a Chern connection $D$.
Then $f^{*}D$ is a Chern connection of $f^{*}b$ on $f^{*}E \to X$.
\end{prop}

\begin{proof}
If $s$ and $t$ are holomorphic sections of $E$ and $\xi$ a tangent vector at a point on $X$ we have
\[
\partial_{\xi} f^{*} b(f^{*}s, \overline{f^{*}t})
= f^{*}(\partial_{f_{*}\xi} b(s, \overline t))
= f^{*}(b(D_{f_{*}\xi}s, \overline{t}))
= f^{*}b(f^{*}D_{\xi} f^{*}s, \overline{f^{*}t}).
\]
As $f^{*}E$ is generated by sections of the form $f^{*}s$, this completes the proof.
\end{proof}


\subsection*{A word on notation}


In a lot of situations we care about, we now only have equality up to things that take values in the kernel of a Hermitian form. For example, if $f : E \to F$ is a Hermitian morphism, it admits an adjoint $f^\dagger$ and we have
\[
b_E(f^\dagger f s, \ov t)
= b_F(f s, \ov{f t})
= b_E(s, \ov t)
\]
for all sections $s$ and $t$,
so $f^\dagger f - \id_E \in \Hom(E, \Ker b_E)$. When working on vector spaces we could deal with this by quotienting out the kernel. If we try the same trick here, we run into the fact that the rank of the sheaf $\Ker b \to X$ may not be constant. We want to do differential geometry to the sections of this sheaf, so this is a problem.

This difficulty is a mirage as it can be taken care of by notation. We define an equivalence relation on sections by saying that two sections are equivalent if their difference is in the kernel of a Hermitian form of interest:


\begin{defi}
If $(E, b) \to X$ is a holomorphic vector bundle with a Hermitian form $b$, and $s, t \in \cc A^k(E)$, we write
\(
s \sim t
\)
if $s - t \in \cc A^k(\Ker b)$.
\end{defi}

In general it's problematic to define differential forms with values in a sheaf, but here it's hopefully clear what $\cc A^k(\Ker b)$ means as $\Ker b$ is a subsheaf of the sheaf of sections of a vector bundle.

Using this notation we can rephrase some of our earlier results on degenerate forms:
\begin{itemize}
\item
If $f : E \to F$ is a Hermitian morphism and $f^\dagger$ an adjoint, then $f^\dagger f \sim \id_E$.

\item
If $f : E \to F$ is a morphism and $f_1^\dagger$ and $f_2^\dagger$ are adjoints of $f$, then $f_1^\dagger \sim f_2^\dagger$.

\item
If $f : E \to F$ is a morphism that admits an adjoint, then $D_{\Hom(E,F)}$ has adjoints $(D_{\Hom(E,F)}f)^\dagger \sim D_{\Hom(F,E)}f^\dagger$.

\item
If $D$ is a Chern connection of $(E,b)$ and $s \sim t$ then $Ds \sim Dt$.

\item
If $D_1$ and $D_2$ are Chern connections of $(E,b)$, then $D_1s \sim D_2s$ for all sections $s$.
\end{itemize}
As adjoints and connections behave well with respect to this relation, we can prove the result we want by essentially the same arguments as in the case where the Hermitian forms are positive-definite; we simply replace ``$=$'' in those proofs by ``$\sim$'' and note that everything else works as before.

An example of this notation's usefulness is the following result, which is annoying to state and use without it. For Hermitian metrics, the $(2,0)$-part of the curvature form is zero. For our Chern connections, the same is not necessarily true; we can't say what happens in the kernel of the form.

\begin{prop}
Let $(E, b) \to X$ be a Hermitian bundle, and let $D$ be a Chern connection for $b$. Then $(D')^2 \sim 0$.
\end{prop}

\begin{proof}
We have
\[
\partial b(s, \ov t)
= b(D's, \ov t) + b(s, \ov{\bar\partial t})
\]
so
\[
0 = \partial^2 b(s, \ov t)
= b((D')^2s, \ov t)
- b(D's, \ov{\bar\partial t})
+ b(D's, \ov{\bar\partial t})
+ b(s, \ov{\bar\partial^2 t})
= b((D')^2s, \ov t)
\]
for all sections $s$ and $t$.
\end{proof}




\subsection*{Sub- and quotient bundles}

Let
\[
\begin{tikzcd}
0 \ar[r] &
S \ar[r,"j"] &
E \ar[r,"q"] &
Q \ar[r] &
0
\end{tikzcd}
\]
be a short exact sequence of holomorphic bundles over $X$, and let $b_E$ be a smooth Hermitian form on $E$. It induces Hermitian forms $b_S$ and $b_Q$ on $S$ and $Q$ as we have seen. Both morphisms $j$ and $q$ also admit adjoints $j^\dagger: E \to S$ and $q^\dagger : Q \to E$. We denote Chern connections of $S$, $E$ and $Q$ by $D_S$, $D_E$ and $D_Q$.




\begin{defi}
The \emph{second fundamental form} of $S$ in $E$ is
\[
\snd(s) := q(D_E(js) - jD_S(s)).
\]
\end{defi}

\begin{prop}
The second fundamental form is an element of $\cc A^{1,0}(\Hom(S,Q))$.
We have
\[
\snd(s)
= q(D_{\Hom(S,E)}j(s))
= q(D_E(js))
\]
for sections $s$ of $S$.
\end{prop}

\begin{proof}
As $qj = 0$ it is clear that $b(s) = q(D_E(js))$. We also have
\[
D_E(js) = D_{\Hom(S,E)}(j)(s) + j(D_Ss),
\]
so $b(s) = q(D_{\Hom(S,E)}(j)(s))$.

As $j$ is holomorphic, we have $D_{\Hom(S,E)}j = D'_{\Hom(S,E)}j$ so $\snd$ has no $(0,1)$-part. It is also clearly $\cc C^\infty$-linear in its tensor field variable. If $f$ is a smooth function, we have
\[
D_E(j(fs))
= df \otimes js + f D_E(js)
\]
so $\snd(fs) = q(D_E(j(fs))) = q(f D_E(js)) = f\snd(s)$.
\end{proof}


\begin{prop}
The second fundamental form admits an adjoint.
\end{prop}

\begin{proof}
We have to show that $\snd(s) \in \Ker b_Q$ when $s$ is a section of $\Ker b_S$. If $s$ is such a section, then $js$ is a section of $\Ker b_E$, so $D_E(js)$ is a section of $\Ker b_E$. Then $q(D_E(js))$ is a section of $\Ker b_Q$ by the construction of $b_Q$.
\end{proof}



Our goal is to calculate the curvature forms of the sub- and quotient bundles. The first stop on the way is a wildly useful list of formulas from Demailly~{{\cite[Theorem~14.3]{demailly-complex}}}.


\begin{prop}
\label{prop:seq-formulas}
\begin{alignat*}{2}
D'_{\Hom(S,E)}j &\sim q^\dagger \circ \snd,
&
\qquad
\bar\partial j &= 0,
\\
D'_{\Hom(E, Q)} q &\sim - \snd \circ j^\dagger,
&
\bar\partial q &= 0,
\\
D'_{\Hom(E,S)} j^\dagger &\sim 0,
&
\bar\partial j^\dagger &\sim b^\dagger \circ q,
\\
D'_{\Hom(Q,E)} q^\dagger &\sim 0,
&
\bar\partial q^\dagger &\sim - j \circ \snd^\dagger,
\\
D'_{\Hom(S,Q)} \snd &\sim 0,
&
\bar\partial \snd^\dagger &\sim 0.
\end{alignat*}
\end{prop}

\begin{proof}
Let $s$ be a section of $S$. Then
\[
j(D_S s) + q^\dagger \snd(s)
\sim D_E(js)
= D_{\Hom(S,E)}j (s) + j(D_S s),
\]
where the first relation is by definition of $D_S$ and $\snd$, so
\[
D_{\Hom(S,E)}j \sim q^\dagger \circ \snd.
\]
The morphism $j$ is holomorphic, so $\bar\partial j = 0$. This proves the first line. The third line follows from the first by taking adjoints.


For the second line, the morphism $q$ is also holomorphic, so $\bar\partial q = 0$. Taking adjoints, we get that $D'_{\Hom(Q,E)} q^\dagger \sim 0$.
We have $\id_E \sim j \circ j^\dagger + q^\dagger \circ q$, so
\begin{align*}
0
&= D'_{\End E} \id_E
\\
&\sim D'_{\Hom(S,E)} j \circ j^\dagger + j \circ D'_{\Hom(E,S)}j^\dagger
+ D'_{\Hom(Q,E)}q^\dagger \circ q + q^\dagger D'_{\Hom(E,Q)} q
\\
&\sim q^\dagger \circ \snd \circ j^\dagger + q^\dagger \circ D'_{\Hom(E,Q)}q.
\end{align*}
Applying $q$, we conclude that $D'_{\Hom(E,Q)}q \sim - \snd \circ j^\dagger$. This proves the second line. Taking adjoints we get the last piece of the fourth line.

Finally we note that
\begin{align*}
D'_{\Hom(S,Q)} \snd
&\sim D'_{\Hom(S,Q)} (q D'_{\Hom(S,E)}j)
\\
&\sim - \snd \circ j^\dagger \circ q^\dagger \circ \snd
+ q^\dagger (D'_{\Hom(S,Q)})^2 j
\sim 0
\end{align*}
as we're dealing with curvature tensors of Hermitian forms which are of type $(1,1)$ up to things that are $\sim 0$. The statement about $\bar\partial \snd^\dagger$ follows by taking adjoints.
\end{proof}



\begin{prop}
Under the smooth splitting $E \to S \oplus Q$ defined by $s \mapsto j^\dagger s \oplus qs$ the Chern connection and curvature form of $E$ are
\begin{align*}
D_E s &\sim
\begin{pmatrix}
j^\dagger & q
\end{pmatrix}^\dagger
\begin{pmatrix}
D_S & - \snd^\dagger
\\
\snd & D_Q
\end{pmatrix}
\begin{pmatrix}
j^\dagger \\ q
\end{pmatrix}
(s),
\\
D_E^2 s &\sim
\begin{pmatrix}
j^\dagger & q
\end{pmatrix}^\dagger
\begin{pmatrix}
D^2_S - \snd^\dagger \wedge \snd & -D'_{\Hom(Q,S)} \snd^\dagger
\\
\bar\partial \snd & D^2_Q - \snd \wedge \snd^\dagger
\end{pmatrix}
\begin{pmatrix}
  j^\dagger \\ q
\end{pmatrix}(s).
\end{align*}
\end{prop}


\begin{proof}
Let $s$ be a section of $E$. We write
\[
s \sim j (j^\dagger s) + q^\dagger( qs).
\]
Then
\[
D_E s
\sim q^\dagger \circ \snd (j^\dagger s)
+ j \circ D_S( j^\dagger s)
- j \circ \snd^\dagger (qs)
+ q^\dagger \circ D_Q (qs).
\]
This proves the first statement.
For each of the terms here, we have
\begin{align*}
D_E(q^\dagger \circ \snd (j^\dagger s))
&\sim -j \circ \snd^\dagger \circ \snd (j^\dagger s)
+ q^\dagger \circ \bar\partial \snd (j^\dagger s)
- q^\dagger \circ \snd \circ D_S(j^\dagger s),
\\
D_E(j \circ D_S( j^\dagger s))
&\sim q^\dagger \circ \snd \circ D_S(j^\dagger s)
+ j \circ D_S^2 (j^\dagger s),
\\
D_E(-j \circ \snd^\dagger (qs))
&\sim - q^\dagger \circ \snd \circ \snd^\dagger (qs)
- j \circ D'_{\Hom(Q,S)}\snd^\dagger (qs)
+ j \circ \snd^\dagger \circ D_Q(qs),
\\
D_E(q^\dagger \circ D_Q (qs))
&\sim -j \circ \snd^\dagger \circ D_Q(qs)
+ q^\dagger \circ D_Q^2 (qs).
\end{align*}
Grouping these together by pre- and postfix morphisms, we get
\begin{align*}
D_E^2 s
&\sim j \bigl( D_S^2 - \snd^\dagger \circ \snd \bigr) j^\dagger s
- j \bigl( D'_{\Hom(Q,S)} \snd^\dagger \bigr) qs
\\
&\qquad
+ q^\dagger \bigl( \bar\partial \snd \bigr) j^\dagger s
+ q^\dagger \bigl( D^2_Q - \snd \circ \snd^\dagger \bigr) qs
\\
&=
\begin{pmatrix}
j^\dagger & q
\end{pmatrix}^\dagger
\begin{pmatrix}
D^2_S - \snd^\dagger \wedge \snd & -D'_{\Hom(Q,S)} \snd^\dagger
\\
\bar\partial \snd & D^2_Q - \snd \wedge \snd^\dagger
\end{pmatrix}
\begin{pmatrix}
  j^\dagger \\ q
\end{pmatrix}(s).
\end{align*}
\end{proof}




\begin{coro}[Codazzi--Griffiths equations]
The curvature tensors of $S$ and $Q$ satisfy
\begin{align*}
R_S(\alpha, \ov\beta, s, \ov t)
&= R_E(\alpha, \ov\beta, js, \ov{jt})
- \tfrac i2 b_Q(\snd(\alpha, s), \ov{\snd(\beta, t)}),
\\
R_Q(\alpha, \ov\beta, s, \ov t)
&= R_E(\alpha, \ov\beta, q^\dagger s, \ov{q^\dagger t})
+ \tfrac i2 b_S(\snd^\dagger(\ov\beta, s), \ov{\snd^\dagger(\ov\alpha, t)}).
\end{align*}
\end{coro}

\begin{proof}
The task here is to convince the reader that our signs are correct and that the arguments go in the right place. Recall that we have a $(1,0)$-form $b$ with values in $\Hom(S,Q)$. Then $b \circ b^\dagger$ and $b^\dagger \circ b$ are both $(1,1)$-forms, with values in $\End Q$ and $\End S$, respectively. Our first claims are that the $(1,1)$-forms
\[
\tfrac i2 h_Q(b \circ b^\dagger(s), \ov s)
\qandq
\tfrac i2 h_S(b^\dagger \circ b(s), \ov s)
\]
are real and seminegative and semipositive for all sections $s$ (of $S$ and $Q$ as appropriate), respectively. For the realness, we have
\begin{align*}
\overline{\rho(\alpha, \ov\beta)}
:= \overline{\tfrac i2 h_Q(b \circ b^\dagger(\alpha, \ov\beta, s), \ov s)}
&= -\tfrac i2 h_Q(s, \ov{b \circ b^\dagger(\alpha,\ov\beta, s)})
\\
&= -\tfrac i2 h_Q(b \circ b^\dagger(\ov\alpha, \beta, s), \ov{s})
\\
&= \tfrac i2 h_Q(b \circ b^\dagger(\beta, \ov\alpha, s), \ov{s})
= \rho(\beta, \ov\alpha),
\end{align*}
so the first form is real. The second is also real by a very similar calculation.

TODO: We'd like to show that
\[
\tfrac i2 b \wedge b^\dagger (\alpha, \ov\beta, s)
= b(\alpha, b^\dagger(\ov\beta, s)),
\]
where the left-hand side is the value of the $(1,1)$-form $\frac i2 b \wedge b^\dagger$. Maybe this needs to be interpreted as ``the $(1,1)$-form is the imaginary part of a Hermitian form with values in $\End Q$, which is the right-hand side''.

Now note that
\[
\tfrac i2 h_S(b \circ b^\dagger(\alpha, \ov\beta, s), \ov t)
= \tfrac i2 h_S(b(\alpha, b^\dagger(\ov \beta, s)), \ov t)
= \tfrac i2 h_S(b^\dagger(\ov \beta, s), \ov{b^\dagger(\ov \alpha, t)}),
\]
so the second form is semipositive. The first form looks very similar, but a careful inspection will reveal that if we were to write it out in local coordinates its terms would be of the form $\frac i2 d\bar z_k \wedge dz_j$. This explains why we have
\begin{align*}
\tfrac i2 h_Q(b^\dagger \circ b(\alpha, \ov\beta, s), \ov t)
&= -\tfrac i2 h_Q(b^\dagger \circ b(\ov\beta, \alpha, s), \ov t)
\\
&= -\tfrac i2 h_Q(b^\dagger(\ov\beta, b(\alpha, s)), \ov t)
= -\tfrac i2 h_Q(b(\alpha, s), \ov{b(\ov\beta, t)}),
\end{align*}
so the form is seminegative.
\end{proof}



\begin{coro}
Let $E \to X$ be a holomorphic vector bundle with two Hermitian forms $b_1$ and $b_2$. Assume that $h := b_1 + b_2$ is positive-definite. The curvature tensor of $h$ is
$$
\displaylines{
R_h(\alpha, \ov\beta, s, \ov t)
= R_{b_1}(\alpha, \ov\beta, s, \ov t)
+ R_{b_2}(\alpha, \ov\beta, s, \ov t)
\hfill\cr\hfill{}
- b_1(h^{-1}b_2((D_{b_1,\alpha} - D_{b_2,\alpha})s), \ov{h^{-1}b_2(D_{b_1,\beta} - D_{b_2,\beta})t})
\phantom{.}
\cr\hfill{}
- b_2(h^{-1}b_1((D_{b_1,\alpha} - D_{b_2,\alpha})s), \ov{h^{-1}b_1(D_{b_1,\beta} - D_{b_2,\beta})t}).
}
$$
\end{coro}




\section{Grassmannian bundles}
\label{sec:grassmannian-bundles}

TODO.



\section{Complexified K\"ahler cone bundles}
\label{sec:compl-kahl-cone}

TODO.




\section{Curvature tensors as Hermitian forms}
\label{sec:curvature-tensors-as}

Let $(E,h) \to X$ be a holomorphic vector bundle with a Hermitian metric $h$. The curvature tensor of $E$ is
\[
R(\alpha, \ov\beta, s, \ov t)
= h(\tfrac i2 \Theta_{\alpha\ov\beta} s, \ov t).
\]
It defines a Hermitian form $b$ on $E \otimes T_X$ that is in general not positive-definite.

TODO: Calculate a Chern connection and a curvature tensor.

Suppose all the sections involved are holomorphic.
We have
\begin{align*}
b(D'(\alpha \otimes s), \ov{\beta \otimes t})
&= \partial h(\tfrac i2 \Theta_{\alpha\ov\beta} s, \ov t)
\\
&= h(\tfrac i2 D'(\Theta_{\alpha\ov\beta} s), \ov t)
\\
&= h(\tfrac i2 (D'_{\End E}\Theta_{\alpha\ov\beta}) s, \ov t)
+ h(\tfrac i2 \Theta_{\alpha\ov\beta} D' s, \ov t)
\\
&= h(\tfrac i2 (D'_{\End E}\Theta_{\alpha\ov\beta}) s, \ov t)
+ R(\alpha, \ov \beta, Ds, \ov t).
\end{align*}
It's tricky to split this up in such a way as to still have something well-defined. Recall that we must respect $f(\beta \otimes t) = f \beta \otimes t = \beta \otimes ft$. Maybe we don't try to split this up? Can we massage
\[
b(D'(\alpha \otimes s), \ov{\beta \otimes t})
= h(\tfrac i2 D'(\Theta_{\alpha\ov\beta} s), \ov t)
\]
into an equation involving the curvature form of $b$?




\bibliographystyle{plain}
\bibliography{sgct}

\end{document}
