\documentclass[10pt,a4paper]{amsart}

\usepackage{lmodern}
\linespread{1.1}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{fancyref}
\usepackage[colorlinks=true]{hyperref}
\usepackage{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{tikz-cd}

\newtheorem{theo}{Theorem}
\newtheorem{prop}[theo]{Proposition}
\newtheorem{lemm}[theo]{Lemma}
\newtheorem{coro}[theo]{Corollary}
\theoremstyle{definition}
\newtheorem{exam}[theo]{Example}

\def\ov#1{\overline{#1}}
\def\qandq{\quad\text{and}\quad}
\newcommand{\kk}[1]{\mathbb{#1}}
\newcommand{\cc}[1]{\mathcal{#1}}

\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\im}{Im}
\DeclareMathOperator{\id}{Id}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Vol}{Vol}
\DeclareMathOperator{\End}{End}

\allowdisplaybreaks

\author{Gunnar \TH\'or Magn\'usson}
\address{Hafnarfjörður, Iceland}
\email{gunnar@magnusson.io}
\date{\today}
\title{Examples of K\"ahler curvature tensors}

\begin{document}

\begin{abstract}
Let $R$ be the curvature tensor of a K\"ahler metric, and let $H$, $r$ and $s$
be its holomorphic sectional, Ricci and scalar curvatures. We show that
$$
\frac{1}{\Vol S(V)} \int_{S(V)}
\!\!\!\!
H(v)^2 \, d\mu(v)
= \frac{6}{\binom{n+3}{4}} \bigl(
|R|^2 + 2 |r|^2 + s^2
\bigr),
$$
where $S(V)$ is the unit sphere.
We also work out local examples that illustrate the folklore that the reverse
implications between the positivity of the various curvature tensors of
K\"ahler metrics all fail.
\end{abstract}

\maketitle

\section{Introduction}

It is well-known that the curvature properties of K\"ahler metrics ``flow
downward''.
To give meaning to this statement,
let us say that a tensor $\alpha$ \emph{dominates} a tensor $\beta$
if any of the following implications hold:
\begin{align*}
\alpha > 0 &\implies \beta > 0,
&
\alpha &\geq 0 \implies \beta \geq 0,
\\
\alpha < 0 &\implies \beta < 0,
&
\alpha &\leq 0 \implies \beta \leq 0.
\end{align*}
Strictly speaking we need to specify what it means for a tensor to be
``positive'' or ``negative'' for this to make sense.
This is fairly clear for the tensors we'll discuss here, except for the
K\"ahler curvature tensor itself which we mean to be Griffiths positive (or
negative).

If $R$ is the curvature tensor of a K\"ahler metric, its various derived
curvature tensors are the holomorphic sectional curvature $H$, the Ricci
tensor $r$ and the scalar curvature~$s$.
Then
$$
\begin{tikzcd}
& R \ar[dl] \ar[dr] &
\\
H \ar[dr] & & r \ar[dl]
\\
& s &
\end{tikzcd}
$$
where an arrow $\alpha \to \beta$ means that $\alpha$ dominates $\beta$.
This is well-known (see for example \cite{zheng2000complex}) but we give a
quick proof in Proposition~\ref{prop:dominance}.

The witches of the field know that neither $H$ nor $r$ dominates the other.
The best evidence for this seems to be indirect.
We for example know that Hirzebruch surfaces (surfaces that are projective
bundles over $\kk P^1$) carry metrics with positive holomorphic sectional
curvature, but carry no metric of positive Ricci curvature as their anticanonical
bundle is not ample;
see \cite{hitchin1975curvature,alvarez2018projectivized,yang2019hirzebruch}.
There are some suggestive results in the negatively curved case, though,
where a compact K\"ahler manifold that carries a metric of negative holomorphic
sectional curvature also carries a possibly different metric of negative Ricci
curvature; see \cite{wu2016negative,tosatti2017extension,diverio2019quasi}.
Could it be the same metric?
We don't know.
Wouldn't it be nice if some easy calculations said it was?
Sure would.

It is also folklore that all of the reverse relations in the diagram fail (see
again \cite{zheng2000complex}). Some do not hold by general considerations; for
example Hirzebruch surfaces
again carry a metric of positive holomorphic sectional curvature but not one of
positive Griffiths curvature, for otherwise they would be $\kk P^2$ by the
Siu--Yau's solution of the Frankel conjecture~\cite{siu1980compact} and, lo,
they are not.

We feel like it would be nice to have some examples that show explicitly what
doesn't hold here.
In this note we work out local examples that show that the reverse of any known
positivity implication fails.
As other people have noted it would be nice to have compact examples of these
failures but it is fairly difficult to get one's hands on those.
Along the way we discuss some basic properties of algebraic curvature tensors
and give a representation-theoretic derivation of some integrals of the
holomorphic sectional curvature over the unit sphere, which give a new proof
that the holomorphic sectional curvature determines the whole curvature tensor.



\section{Algebraic curvature tensors}



\subsection*{K\"ahler curvature tensors come from symmetric tensors}

Let $V$ be a complex vector space of dimension $n$, which we think of as the
tangent space of a complex manifold at a given point.
The curvature tensor $R$ of a Hermitian metric on the manifold identifies with
a Hermitian form $q$ on $V \otimes V$, defined by
$$
R(x, \ov y, z, \ov w)
= q(x \otimes z, \ov{y \otimes w}).
$$
If the metric is K\"ahler we get an additional symmetry
$R(x, \ov y, z, \ov w) = R(z, \ov y, x, \ov w)$
(and the ones induced by conjugating).
A nice alternate reference for what we discuss here is
\cite{algebraic-kahler-curvature}.

We write $B(V)$ for the real vector space of Hermitian forms on $V$.
The curvature tensor of a Hermitian metric is then just a member of $B(V
\otimes V)$. We call such an element an \emph{algebraic Hermitian curvature
tensor}, and one that satisfies the additional symmetry of a K\"ahler curvature
tensor an \emph{algebraic K\"ahler curvature tensor}.

The decomposition $V \otimes V = \bigwedge^2 V \oplus \Sym^2 V$ is standard.
It implies that a Hermitian form on $V \otimes V$ decomposes into components
$$
q = \begin{pmatrix}
q_{\wedge^2 V} & q_{(\Sym^2V, \wedge^2 V)}
\\
q_{(\wedge^2 V, \Sym^2V)} & q_{\Sym^2 V},
\end{pmatrix}
$$
where $\ov{q}_{(\wedge^2 V, \Sym^2V)}^t = q_{(\Sym^2V, \wedge^2 V)}$.
Denote the symmetrization map by
$$
\Pi_2 : V \otimes V \to \Sym^2 V,
\quad
x \otimes y \mapsto \tfrac 12 (x \otimes y + y \otimes x).
$$
It is a surjective linear morphism that realizes the space of symmetric tensors
as a subspace of $V \otimes V$.
The usual definition of that space is as a quotient of $V \otimes V$ by the
ideal generated by $x \otimes y - y \otimes x$.
As the field $\kk C$ has characteristic zero these spaces are isomorphic,
so the difference between them isn't very important to us.


\begin{prop}
A tensor $R \in B(V \otimes V)$ is an algebraic K\"ahler curvature tensor
if and only if there exists an element $\hat R \in B(\Sym^2 V)$ such that $R =
\Pi_2^* \hat R$.
\end{prop}

\begin{proof}
Suppose $R \in B(V \otimes V)$ is an algebraic K\"ahler curvature tensor.
We define
$$
\hat R(x \odot z, \ov{y \odot w})
= R(x, \ov y, z, \ov w),
$$
which is well-defined because $R$ is K\"ahler, and we have $R = \Pi_2^* \hat R$.

Conversely, let $\hat R \in B(\Sym^2 V)$, and define $R = \Pi_2^* \hat R$.
Then
$$
R(x, \ov y, z, \ov w)
= \hat R(x \odot z, \ov{y \odot w})
= \hat R(z \odot x, \ov{w \odot y})
= R(z, \ov y, x, \ov w)
$$
is an algebraic K\"ahler curvature tensor.
\end{proof}

As an aside, this proposition explains why we only ever talk about Griffits
positivity of K\"ahler metrics.
A Hermitian metric is Nakano positive if its curvature tensor is
positive-definite as a Hermitian form on $V \otimes V$.
kernel is never positive-definite.
Therefore a K\"ahler metric is never Nakano positive when $\pi$ is not
injective, which happens when
$n^2 = \dim V \otimes V > \dim \Sym^2 V = \binom {n+1}2$,
that is, when $n > 1$.

There should then be a notion of positivity for K\"ahler curvature tensors that
interpolates between Griffiths positivity and Nakano positivity and is perhaps
more geometrically motivated than Griffiths positivity, where we would say that
such a tensor is positive if its Hermitian form on $\Sym^2 V$ is
positive-definite.
In light of Siu and Yau's theorem any compact K\"ahler manifold that
admits a metric so positive is just projective space, so we don't need to think
about it anymore.



\subsection*{Holomorphic sectional curvature determines the whole tensor}

It is well-known that the holomorphic sectional curvature of a K\"ahler metric
determines the whole curvature tensor (\cite{zheng2000complex}, again) and that
its average over the unit sphere is a multiple of the scalar curvature~
\cite{berger1965varietes,zheng2000complex,diverio2020kobayashi}.
The proofs of this in the litarature are somewhat mysterious;
the first proceeds by algebraic manipulations that are not explained or
motivated and only seem justified post-hoc by virtue of giving the right answer;
the second is by a brute force integration that seems to imply the result is
due to the seeming coincidence that integrating $|z_j^4|$ and $|z_j|^2 |z_k|^2$
over the sphere gives the same result up to a factor of 2.
We intend to show that both results flow from the same representation-theoretic
source.


Our starting point there involves vector-valued integration, so let's recall
some basic facts.
Let $V$ and $W$ be finite-dimensional complex vector spaces equipped with their
Lebesgue measures.
If $f : V \to W$ is a continuous function and $X \subset V$ a measurable subset
we define $\int_X f(v) d\mu(v) \in W$ to be the vector we get after picking
bases and integrating coordinate by coordinate.
If $L : W \to Z$ is a linear map then $L \int_X f(v) d\mu(v) = \int_X Lf(v)
d\mu(v)$, because that holds at every step of the definition of the integral of
a measurable function.
This implies the integral is well-defined, as it behaves correctly with respect
to a change of bases.
It also implies that if $f : V \to V$ is continuous then the trace commutes
with the integral.

Let's suppose $V$ is equipped with a Hermitian inner product $h$
and let's write $d\mu$ for the induced volume form on the unit sphere $S(V)$ in
$V$. For $v \in V$ we define $v \otimes v^*$ to be the linear map
$h^*(\ov v) v = x \mapsto h(x, \ov v) v$.
Note that if $f \in \End V$ then $f \circ v \otimes v^* = h^*(\ov v) f(v)$
and thus $\tr(f \circ v \otimes v^*) = h(fv, \ov v)$.
The following is known to quantum information theorists; see
\cite{harrow2013church} and \cite[Chapter~7]{watrous2018theory}.

\begin{prop}
\label{prop:int-id}
Denote by $\Pi_d : V^{\otimes d} \to V^{\otimes d}$ the projection onto $\Sym^d V$.
Then
$$
\frac{1}{\Vol S(V)}
\int_{S(V)} (v \otimes v^*)^{d} d\mu(v)
= \frac{1}{\binom{n+d-1}{d}} \Pi_d.
$$
\end{prop}

\begin{proof}
Let's denote the map defined by the integral by $L$. If $g \in \End V$ is
unitary with respect to $h$ we have $h(g x, \ov v) = h(x, \ov{g^{-1} v})$ and
$|\!\det g| = 1$ so the change of basis formula implies that $L g^d = g^d L$.
The map $L$ is then an interleaving operator of the representation $V^{\otimes d}$
of the unitary group.

The integral is thus a sum of multiples of the projections onto the irreducible
factors of $V^{\otimes d}$.
Note however that the integral takes values in $\Sym^d V$ because the integrand
is invariant under $S_d$, so $L = \lambda \Pi_d$ for some scalar $\lambda$.
For a unit vector $v$ we have $\tr v \otimes v^* = |v|^2 = 1$.
Taking the trace we then find that $1 = \lambda \binom{n+d-1}{d}$,
which implies the result.
\end{proof}


\begin{coro}
\begin{align*}
\frac{1}{\Vol S(V)} \int_{S(V)}
\!\!\!
H(v) \, d\mu(v)
&= \frac{1}{\binom{n+1}{2}} s.
\\
\frac{1}{\Vol S(V)} \int_{S(V)}
\!\!\!
H(v)^2 \, d\mu(v)
&=
\frac{6}{\binom{n+3}4} (|R|^2 + 2|r|^2 + s^2).
\end{align*}
\end{coro}


\begin{proof}
If $f \in \End V^{\otimes 2}$ is symmetric then taking the trace of the
equation in Proposition~\ref{prop:int-id} for $d = 2$ gives
$$
\frac{1}{\Vol S(V)}
\int_{S(V)} h^{\otimes 2}(f(v \otimes v), \ov{v \otimes v}) \, d\mu(v)
= \frac{1}{\binom{n+1}{2}} \tr f.
$$
If $q \in B(\Sym^2)$ is the Hermitian form defined by the curvature tensor $R$,
then applying this to $f = (h^{\otimes 2})^{-1} \Pi_2^* b$ gives the first
equation.

For the second equation, we consider $f \otimes f \in \End V^{\otimes 4}$ where
$f$ is as before and get that
$$
\frac{1}{\Vol S(V)}
\int_{S(V)} h^{\otimes 2}(f(v \otimes v), \ov{v \otimes v})^2 \, d\mu(v)
= \frac{1}{\binom{n+3}{4}}
\tr (f \circ \Pi_4).
$$
We know that $\Pi_d = \sum_{\sigma \in S_d} W_\sigma$, where $S_d$ is the
symmetric group on $d$ letters and $W_\sigma(v_1 \otimes \cdots \otimes v_d) =
v_{\sigma(1)} \otimes \cdots \otimes v_{\sigma(d)}$.
We note that $W_{\sigma\rho} = W_\sigma W_\rho$.
The group $S_4$ is generated by
$\sigma = (1 \  2)$ and
$\rho = (1 \  2 \  3 \  4)$, explicitly as%
\def\ze#1{{\color{red}#1}}
\def\on#1{{\color{blue}#1}}
\def\tw#1{{\color{brown}#1}}
\def\th#1{{\color{violet}#1}}
\begin{align*}
\ze{(1\;2\;3\;4)} &= e
	     &
\on{(2\;3\;4\;1)} &= \rho
	     &
\tw{(3\;4\;1\;2)} &= \rho \rho
	     &
\th{(4\;1\;2\;3)} &= \rho \rho \rho
\\
\ze{(2\;1\;3\;4)} &= \sigma
	     &
\on{(1\;3\;4\;2)} &= \rho\sigma
	     &
\tw{(2\;4\;1\;3)} &= \rho \sigma \rho
	     &
\th{(4\;2\;1\;3)} &= \rho \rho \rho\sigma
\\
\ze{(1\;3\;2\;4)} &= \rho \rho \rho \sigma \rho
	     &
\on{(3\;2\;4\;1)} &= \sigma\rho
	     &
\tw{(4\;3\;1\;2)} &= \sigma\rho \rho
	     &
\th{(4\;1\;3\;2)} &= \rho \rho \sigma \rho
\\
\ze{(1\;2\;4\;3)} &= \rho \rho \sigma \rho \rho
	     &
\on{(3\;1\;4\;2)} &= \sigma \rho \sigma
	     &
\tw{(3\;4\;2\;1)} &= \rho \rho\sigma
	     &
\th{(3\;1\;2\;4)} &= \rho \sigma \rho \rho
\\
\ze{(2\;1\;4\;3)} &= \rho \rho \sigma \rho \rho \sigma
		  &
\th{(3\;2\;1\;4)} &= \rho \sigma \rho \rho \sigma
	     &
\tw{(1\;4\;2\;3)} &=  \rho\sigma \rho \sigma
	     &
\th{(1\;4\;3\;2)} &= \sigma \rho \rho \sigma \rho
\\
\ze{(2\;3\;1\;4)} &= \rho \rho \rho \sigma \rho \sigma
	     &
\th{(2\;4\;3\;1)} &= \sigma \rho \rho \sigma \rho \sigma
	     &
\tw{(4\;3\;2\;1)} &= \sigma \rho \rho \sigma
	     &
\th{(4\;2\;3\;1)} &= \rho \rho \sigma \rho\sigma
\end{align*}
as the reader can have fun verifying (or write a program to generate these, as
I did).
We have colored the elements of $S_4$ by the number of times $\rho$ appears
in their expressions modulo 4, foreshadowingly.
As $f$ is symmetric we have $f \otimes f \circ W_{\sigma} = f \otimes f$
so we only need to figure out what the traces of $f \otimes f \circ W_{\rho^k}$
are for $k = 1, 2, 3$ because even I can tell that
$\tr(f \otimes f \circ W_{e}) = (\tr f)^2$.

To do this we pick a basis $(e_j)$ of $V$ and denote by $(e_j
\otimes e_k \otimes e_l \otimes e_m)$ the induced basis of $V^{\otimes 4}$.
Denote by $f_{jk,lm}$ the coefficients of the matrix of $f$ in the
corresponding basis $(e_j \otimes e_k)$ of $V^{\otimes 2}$, so the matrix of
$f \otimes f$ has the coefficients $f_{jk,lm} f_{pq,rs}$.
Using Einstein notation we then have
\begin{align*}
\tr(f \otimes f \circ W_{\rho})
&= f_{kp,jk} f_{qj,pq}
\\
&= (\tr_{14} f)_{pj} (\tr_{14} f)_{jp}
= \ov{(\tr_{14} f)_{jp}} (\tr_{14} f)_{jp}
= |\! \tr_{14} f|^2,
\\
\tr(f \otimes f \circ W_{\rho^2})
&= f_{pq,jk} f_{jk,pq}
= \ov f_{\! jk,pq} f_{jk,pq}
= |f|^2,
\\
\tr(f \otimes f \circ W_{\rho^3})
&= f_{qj,jk} f_{kp,pq}
\\
&= (\tr_{23} f)_{qk} (\tr_{23} f)_{kq}
= \ov{(\tr_{23} f)_{kq}} (\tr_{23} f)_{kq}
= |\! \tr_{23} f|^2,
\end{align*}
where $\tr_{jk} : \End V^{\otimes 4} \to \End V^{\otimes 2}$ is the partial
trace in $j$ and $k$ and the equalities involving conjugations hold because
$f^\dagger = f$. As $f$ is symmetric the partial traces $\tr_{jk} f$ are all
equal to $h^{-1} r$.

Looking at the generator table above, we see that after discarding $\sigma$
the term $|f|^2 = |R|^2$ appears 6 times, as does the term $(\tr f)^2 = s^2$,
and that $|\!\tr_{jk} f|^2 = |r|^2$ appears $4 + 8 = 2 \cdot 6$ times.
\end{proof}

If we just have $f \in \End V^{\otimes 2}$ with $f$ Hermitian (that is, $f$ comes
from a Hermitian curvature tensor $R$) we get the 24 traces
\begin{align*}
\tr(f \otimes f \circ W_{(jklm)})
&= f_{jk,jk} f_{lm,lm}
= \ov{\tr \tr_{13} f} \tr \tr_{13} f
= (\tr f)^2,
\\
\tr(f \otimes f \circ W_{(jkml)})
&= f_{jk,jk} f_{ml,lm}
= \ov{\tr \tr_{13} f} \tr \tr_{14} f
= \tr(\tr_{14} f) \, \tr f,
\\
\tr(f \otimes f \circ W_{(jlkm)})
&= f_{jl,jk} f_{km,lm}
= \ov{(\tr_{13} f)_{kl}} (\tr_{24} f)_{kl}
= \langle \tr_{24} f, \ov{\tr_{13} f} \rangle,
\\
\tr(f \otimes f \circ W_{(jlmk)})
&= f_{jl,jk} f_{mk,lm}
= \ov{(\tr_{13} f)_{kl}} (\tr_{14} f)_{kl}
= \langle \tr_{14} f, \ov{\tr_{13} f} \rangle,
\\
\tr(f \otimes f \circ W_{(jmkl)})
&= f_{jm,jk} f_{kl,lm}
= \ov{(\tr_{13} f)_{km}} (\tr_{23} f)_{km}
= \langle \tr_{23} f, \ov{\tr_{13} f} \rangle,
\\
\tr(f \otimes f \circ W_{(jmlk)})
&= f_{jm,jk} f_{lk,lm}
= \ov{(\tr_{13} f)_{km}} (\tr_{13} f)_{km}
= |\!\tr_{13} f|^2,
\\
\tr(f \otimes f \circ W_{(kjlm)})
&= f_{kj,jk} f_{lm,lm}
= \ov{\tr \tr_{14} f} \tr \tr_{13} f
= \tr(\tr_{14}f) \, \tr f,
\\
\tr(f \otimes f \circ W_{(kjml)})
&= f_{kj,jk} f_{ml,lm}
= \ov{\tr \tr_{14} f} \tr \tr_{14} f
= \tr(\tr_{14} f)^2,
\\
\tr(f \otimes f \circ W_{(kljm)})
&= f_{kl,jk} f_{jm,lm}
= \ov{(\tr_{23} f)_{jl}} (\tr_{24} f)_{jl}
= \langle \tr_{24} f, \ov{\tr_{23} f} \rangle,
\\
\tr(f \otimes f \circ W_{(klmj)})
&= f_{kl,jk} f_{mj,lm}
= \ov{(\tr_{23} f)_{jl}} (\tr_{14} f)_{jl}
= \langle \tr_{14} f, \ov{\tr_{23} f} \rangle,
\\
\tr(f \otimes f \circ W_{(kmjl)})
&= f_{km,jk} f_{jl,lm}
= \ov{(\tr_{23} f)_{jm}} (\tr_{23} f)_{jm}
= |\! \tr_{23} f|^2,
\\
\tr(f \otimes f \circ W_{(kmlj)})
&= f_{km,jk} f_{lj,lm}
= \ov{(\tr_{23} f)_{jm}} (\tr_{13} f)_{jm}
= \langle \tr_{13} f, \ov{\tr_{23} f} \rangle,
\\
\tr(f \otimes f \circ W_{(ljkm)})
&= f_{lj,jk} f_{km,lm}
= \ov{(\tr_{14} f)_{kl}} (\tr_{24} f)_{kl}
= \langle \tr_{24} f, \ov{\tr_{14} f} \rangle,
\\
\tr(f \otimes f \circ W_{(ljmk)})
&= f_{lj,jk} f_{mk,lm}
= \ov{(\tr_{14} f)_{kl}} (\tr_{14} f)_{kl}
= |\! \tr_{14} f|^2,
\\
\tr(f \otimes f \circ W_{(lkjm)})
&= f_{lk,jk} f_{jm,lm}
= \ov{(\tr_{24} f)_{jl}} (\tr_{24} f)_{jl}
= |\! \tr_{24} f|^2,
\\
\tr(f \otimes f \circ W_{(lkmj)})
&= f_{lk,jk} f_{mj,lm}
= \ov{(\tr_{24} f)_{jl}} (\tr_{14} f)_{jl}
= \langle \tr_{14} f, \ov{\tr_{24} f} \rangle,
\\
\tr(f \otimes f \circ W_{(lmjk)})
&= f_{lm,jk} f_{jk,lm}
= \ov{f_{jk,lm}} f_{jk,lm}
= |f|^2,
\\
\tr(f \otimes f \circ W_{(lmkj)})
&= f_{lm,jk} f_{kj,lm}
= \ov{f_{jk,lm}} f_{kj,lm}
= \langle f \circ W_{(12)}, \ov f \rangle ,
\\
\tr(f \otimes f \circ W_{(mjkl)})
&= f_{mj,jk} f_{kl,lm}
= \ov{(\tr_{14} f)_{km}} (\tr_{23} f)_{km}
= \langle \tr_{23} f, \ov{\tr_{14} f} \rangle,
\\
\tr(f \otimes f \circ W_{(mjlk)})
&= f_{mj,jk} f_{lk,lm}
= \ov{(\tr_{14} f)_{km}} (\tr_{13} f)_{km}
= \langle \tr_{13} f, \ov{\tr_{14} f} \rangle,
\\
\tr(f \otimes f \circ W_{(mkjl)})
&= f_{mk,jk} f_{jl,lm}
= \ov{(\tr_{24} f)_{jm}} (\tr_{23} f)_{jm}
= \langle \tr_{23} f, \ov{\tr_{24} f} \rangle,
\\
\tr(f \otimes f \circ W_{(mklj)})
&= f_{mk,jk} f_{lj,lm}
= \ov{(\tr_{24} f)_{jm}} (\tr_{13} f)_{jm}
= \langle \tr_{13} f, \ov{\tr_{24} f} \rangle,
\\
\tr(f \otimes f \circ W_{(mljk)})
&= f_{ml,jk} f_{jk,lm}
= \ov{f_{jk,ml}} f_{jk,lm}
= \langle f, \ov{W_{(12)} \circ f} \rangle,
\\
\tr(f \otimes f \circ W_{(mlkj)})
&= f_{ml,jk} f_{kj,lm}
= \ov{f_{jk,ml}} f_{kj,lm}
= \langle f \circ W_{(12)}, \ov{W_{(12)} \circ f} \rangle.
\end{align*}
The sum of the 16 different $\langle \tr_{jk} f, \ov{\tr_{ml} f} \rangle$
factors is $|\!\tr_{13} f + \tr_{14} f + \tr_{23} f + \tr_{24} f|^2$.
Note that as $f$ is Hermitian we have
$$
(\tr_{13} f)^\dagger = \tr_{13} f,
\quad
(\tr_{14} f)^\dagger = \tr_{23} f,
\quad
(\tr_{24} f)^\dagger = \tr_{24} f.
$$
The sum of the 4 different total trace factors is likewise
$(\tr f + \tr(\tr_{14} f))^2$ (as the total traces as real because $f$ is
Hermitian).

\def\<{\langle}
\def\>{\rangle}
Write $\sigma = W_{(12)} \in \End V^{\otimes 2}$.
Clearly $\sigma^2 = \id$ so its eigenvalues are $1$ and $-1$, and in fact
the splitting $V^{\otimes 2} = \Sym^2 V \oplus \bigwedge^2 V$
is according to the eigenspaces of $\sigma$,
which acts as $\id$ on $\Sym^2V$ and $-\id$ on $\bigwedge^2 V$.
Writing $\sigma$ out according to this splitting we have
$$
\sigma = \begin{pmatrix}
\id & 0
\\
0 & -\id
\end{pmatrix}
$$
from which it is clear that $\sigma^\dagger = \sigma$.
Because of this and $\sigma^2 = \id$ we get
$$
\< f \sigma, \ov g \> = \<f, \ov{g\sigma}\>,
\quad
\< \sigma f , \ov g \> = \<f, \ov{\sigma g}\>,
\quad
\< \sigma f, \ov{\sigma g} \> = \<f, \ov g\>.
$$
Note that $\Pi_2 = \frac12(\id + \sigma)$.
For any $f \in \End V^{\otimes 2}$ its symmetric part is $\Pi_2 \circ f \circ
\Pi_2$ and
$$
\Pi_2 \circ f \circ \Pi_2
= \tfrac14 (\id + \sigma) f (\id + \sigma)
= \tfrac14 (f + f \sigma + \sigma f + \sigma f \sigma).
$$
Then
\begin{align*}
16 |\Pi_2 \circ f \circ \Pi_2|^2
&= |f|^2
+ \< f, f \sigma \>
+ \< f, \sigma f \>
+ \< f, \sigma f \sigma \>
\\
&\qquad
+ \< f \sigma, f \>
+ \< f \sigma, f \sigma \>
+ \< f \sigma, \sigma f \>
+ \< f \sigma, \sigma f \sigma \>
\\
&\qquad
+ \< \sigma f, f \>
+ \< \sigma f, f \sigma \>
+ \< \sigma f, \sigma f \>
+ \< \sigma f, \sigma f \sigma \>
\\
&\qquad
+ \< \sigma f \sigma, f \>
+ \< \sigma f \sigma, f \sigma \>
+ \< \sigma f \sigma, \sigma f \>
+ \< \sigma f \sigma, \sigma f \sigma \>
\\
&= |f|^2
+ \< f \sigma, f \>
+ \< \sigma f, f \>
+ \< \sigma f \sigma, f \>
\\
&\qquad
+ \< f \sigma, f \>
+ |f|^2
+ \< \sigma f \sigma, f \>
+ \< \sigma f , f \>
\\
&\qquad
+ \< \sigma f, f \>
+ \< \sigma f \sigma, f \>
+ |f|^2
+ \< f\sigma , f \>
\\
&\qquad
+ \< \sigma f \sigma, f \>
+ \< \sigma f, f \>
+ \< f \sigma, f \>
+ |f|^2
\\
&= 4(|f|^2
+ \< f \sigma, f \>
+ \< \sigma f, f \>
+ \< \sigma f \sigma, f \>).
\end{align*}
Then finally
$$
\frac{1}{\Vol S(V)}
\int_{S(V)}
\!\!\!
H(v)^2 \, d\mu(v)
= \frac{1}{\binom{n+3}{4}}
\bigl(
4 |R_{\Sym^2 V}|^2
% One pair of these are equal
+ |r_1 + r_2 + r_3 + r_4|^2
+ (s_1 + s_2)^2
\bigr),
$$
where
$R_{\Sym^2 V} = \Pi_2 \circ R \circ \Pi_2$,
$r_j$ are the Ricci tensors of $R$ and $s_j$ are its scalar curvatures.



The equation for the $L^2$-norm of $H$ now gives an alternate reason for why
$H$ determines the whole curvature tensor.
It also shows that knowing a bound on $H$ gives a bound on $R$, which is
usually observed as a corollary of a polarization identity.
The polarization identity shows that if $|H|^2 \leq C$ then $|R|^2 \leq
24 C$ because it involves applying the triangle inequality to a sum of 24 factors.
We get here that $|R|^2 \leq 6\binom{n+3}{4} C$, which is arbitrarily worse!


\subsection*{Dominance}

The following statement is classical but quickly proved now that we're here.

\begin{prop}
\label{prop:dominance}
If $R$ is an algebraic K\"ahler curvature tensor, then
$$
\begin{tikzcd}
& R \ar[dl] \ar[dr] &
\\
H \ar[dr] & & r \ar[dl]
\\
& s &
\end{tikzcd}
$$
\end{prop}

\begin{proof}
We prove the statements under the assumption that the tensors are positive.
The proofs for semipositivity and (semi)negativity are the same.

If $R$ is Griffiths positive, then by definition $R(x, \ov x, y, \ov y) > 0$
for all nonzero $x, y$. Then $R(x, \ov x, x, \ov x) > 0$ so $H(x) > 0$.
The Ricci tensor is the trace
$$
r(x, \ov y) = \sum_{j = 1}^n R(x, \ov y, e_j, \ov e_j)
$$
in an orthonormal basis $(e_j)$, so it is positive-definite.
The scalar curvature is the trace
$$
s = \sum_{j=1}^n r(e_j, \ov e_j)
$$
in an orthonormal basis, so $s > 0$ if $r$ is positive-definite.
As we saw before we have
\[
\frac{1}{\Vol S(V)}
\int_{S(V)} H(v) \, d\mu(v)
= \frac{1}{\binom{n+1}{2}} s
\]
so $s > 0$ if $H > 0$.
\end{proof}



\section{Examples of non-dominance in dimension two}

Let's define
\begin{align*}
\cc H(V)
&= \{ \text{germs of smooth Hermitian metrics on $V$ at $0$} \},
\\
\cc K(V)
&= \{ \text{germs of smooth K\"ahler metrics on $V$ at $0$} \}.
\end{align*}
We just write $\cc H$ if $V$ is clear from context and will usually mean $V =
\kk C^n$. Note that $\cc K$ and $\cc H$ are only convex open cones in real
vector spaces and not full subspaces.
Evaluating the curvature tensor of a metric at $0$ defines nonlinear maps and a
commutative diagram
$$
\begin{tikzcd}
\cc K \ar[d] \ar[r,"\cc R"] & B(\operatorname{Sym}^2 \kk C^n)\ar[d,"\pi^*"]
\\
\cc H \ar[r, "\cc R"] & B(\kk C^n \otimes \kk C^n).
\end{tikzcd}
$$
Both vertical arrows are of course injective.
The horizonal arrows are neither injective nor surjective in general.


Let $H$ be the matrix of a Hermitian metric $h$ in a neighborhood around $0$ in
$V$. The Chern connection of $h$ is given by $D^{1,0} = H^{-1}\partial H$ and its
curvature form is
$$
F = \tfrac i2 \bar\partial(H^{-1}\partial H)
= - \tfrac i2 H^{-1} \bar\partial H \wedge H^{-1} \partial H
- \tfrac i2 H^{-1} \partial \bar\partial H.
$$
The curvature tensor is then $R(x,\ov y, z, \ov w) = h(F(x, \ov y) z, \ov w)$.

We can pick coordinates such that $H = I_n$ at $0$.
If $h$ is K\"ahler we can further get $\partial H = 0$ at $0$.
We will arrange this in our examples.
We now restrict to $V = \kk C^2$.

\begin{prop}
The image of $\cc K(\kk C^2)$ in $B(\Sym^2 \kk C^2)$ is
$$
\left\{
\begin{pmatrix}
a & b & c
\\
\ov b & c & e
\\
c & \ov e & f
\end{pmatrix}
\biggm|
\text{$a, c, f \in \kk R$, $b, e \in \kk C$}
\right\}.
$$
\end{prop}


\begin{proof}
If $h$ is a Hermitian metric on a neighborhood of $0$ in $\kk C^2$
we can write it in matrix form as
$$
H = \begin{pmatrix}
a & c
\\
\ov c & b
\end{pmatrix}.
$$
After a change of coordinates, we can assume that $H(0) = I_2$.
Assuming $h$ is K\"ahler we can also arrange that $\partial H(0) = 0$.
Then the matrix of the curvature tensor in the basis
\def\tg#1#2{\frac{\partial}{\partial #1} \otimes \frac{\partial}{\partial #2}}
$(\tg zz, \tg zw, \tg wz, \tg ww)$
is
$$
R = \begin{pmatrix}
a_{z \ov z} & \ov c_{z \ov z} & a_{z \ov w} & \ov c_{z \ov w}
\\
c_{z \ov z} & b_{z \ov z} & c_{z \ov w} & b_{z \ov w}
\\
a_{w \ov z} & \ov c_{w \ov z} & a_{w \ov w} & \ov c_{w \ov w}
\\
c_{w \ov z} & b_{w \ov z} & c_{w \ov w} & b_{w \ov w}
\end{pmatrix}.
$$
We've written, for example, $a_{z \ov w}$ instead of $\frac{\partial^2 a}{\partial z \partial \ov w}$.
The metric $h$ is K\"ahler, so we have
$$
a_w = c_{z},
\quad
a_{\ov w} = \ov c_{\ov z},
\quad
b_z = \ov c_w,
\quad
b_{\ov z} = c_{\ov w},
$$
and thus
$$
c_{z \ov z} = a_{w \ov z},
\quad
\ov c_{w \ov w} = b_{z \ov w},
\quad
c_{z \ov w} = a_{w \ov w},
\quad
\ov c_{w \ov z} = b_{z \ov z}
$$
Then we get
$$
R = \begin{pmatrix}
a_{z \ov z} & a_{z \ov w} & a_{z \ov w} & a_{w \ov w}
\\
a_{w \ov z} & a_{w \ov w} & a_{w \ov w} & b_{z \ov w}
\\
a_{w \ov z} & a_{w \ov w} & a_{w \ov w} & b_{z \ov w}
\\
a_{w \ov w} & b_{w \ov z} & b_{w \ov z} & b_{w \ov w}
\end{pmatrix},
$$
where some simplifications happen because $a$ is real so $a_{w \ov z} =
a_{z \ov w}$ and we can ping-pong to $b_{z \ov z} = a_{w \ov w}$.
This is the pullback of the Hermitian form
$$
R = \begin{pmatrix}
a_{z \ov z} & a_{z \ov w} & a_{w \ov w}
\\
a_{w \ov z} & a_{w \ov w} & b_{z \ov w}
\\
a_{w \ov w} & b_{w \ov z} & b_{w \ov w}
\end{pmatrix},
$$
on $\Sym^2 \kk C^2$, as expected.
Note that $a_{w \ov w}$ is real, so not every Hermitian form on
$\Sym^2 \kk C^2$ comes from a curvature tensor of a K\"ahler metric.
(Note that $a_{w \ov w} = b_{z \ov z}$ so this matrix is more symmetric in $a$
and $b$ than it appears.)


We will conclude the proof by taking the above Hermitian matrix as given and
constructing a germ of a K\"ahler metric whose curvature tensor agrees with it
at $0$.
Suppose we let
$$
f(z,w)
= a_{z \ov z} z \bar z
+ a_{z \ov w} z \ov w
+ a_{w \ov z} w \ov z
+ a_{w \ov w} w \bar w.
$$
Then $f(0) = 0$ and $df(0) = 0$ and
$$
\operatorname{Hess} f(0)
= \begin{pmatrix}
a_{z \ov z} & a_{z \ov w}
\\
a_{w \ov z} & a_{w \ov w}
\end{pmatrix}.
$$
We also let
$$
g(z,w)
= a_{w \ov w} z \bar z
+ b_{z \ov w} z \ov w
+ b_{w \ov z} w \ov z
+ b_{w \ov w} w \bar w
$$
and
$$
h(z,w)
= h_1 z \bar z
+ h_2 z \ov w
+ h_3 w \ov z
+ h_4 w \bar w.
$$
We want
$$
a_{w \ov z} \ov z + a_{w \ov w} \bar w
= f_w = h_z
= h_1 \ov z + h_2 \ov w,
$$
so we set $h_1 = a_{w \ov z}$ and $h_2 = a_{w \ov w}$.
We also want
$$
a_{w \ov w} \ov z + b_{z \ov w} \ov w
= g_z = \ov h_w
= \ov h_2 \bar z + \ov h_4 \ov w
$$
so $h_4 = b_{z \ov w}$.
There's no constraint on $h_3$ so we just take $h_3 = \ov h_4$.
The germ of the K\"ahler metric
$$
h(z,w)
= \begin{pmatrix}
1 + f(z,w) & h(z,w)
\\
\ov{h(z,w)} & 1 + g(z,w)
\end{pmatrix}
$$
then has the curvature tensor $R$ above.
\end{proof}

Let's look at some special cases. If the tensor looks like
$$
R =
\begin{pmatrix}
a & 0 & c
\\
0 & c & 0
\\
c & 0 & f
\end{pmatrix}
$$
then
\begin{align*}
R(x, \ov x, x, \ov x)
&= (a x_1^2 + c x_2^2) \ov x_1^2
+ 4 c |x_1|^2 |x_2|^2
+ (c x_1^2 + f x_2^2) \ov x_2^2
\\
&= a |x_1|^4 + f |x_2|^4
+ 2 c \bigl(2 |x_1|^2 |x_2|^2 + \Re(x_1^2 \ov x_2^2) \bigr)
\\
r(x, \ov y)
&=
\begin{pmatrix}\ov y_1 & \ov y_2 \end{pmatrix}
\begin{pmatrix}
a + c & 0 \\ 0 & c + f
\end{pmatrix}
\begin{pmatrix} x_1 \\ x_2 \end{pmatrix},
\\
s &= a + 2c + f,
\end{align*}
and the sign of $R(x, \ov x, x, \ov x)$ is the same as the sign of the
holomorphic sectional curvature $H(x)$.

For complex numbers we have $|\Re z| \leq |z|$, so we have
\begin{align*}
a |x_1|^4 + f |x_2|^4 + 2c |x_1|^2 |x_2|^2
&\leq R(x, \ov x, x, \ov x)
\\
&\leq a |x_1|^4 + f |x_2|^4 + 6c |x_1|^2 |x_2|^2.
\end{align*}
If further $x = x_1 = x_2$ then
$$
R(x, \ov x, x, \ov x)
= (a + 6c + f) |x|^4.
$$

\begin{exam}[$s \not\to H$]
We can find $a, c, f$ such that $a + 2c + f > 0$ but $a + 6c + f < 0$,
(like $a = f = 4$, $c = -3$), so the scalar curvature does not dominate
the holomorphic sectional curvature.
\end{exam}

\begin{exam}[$s \not\to r$]
Take $a = 4$, $f = -1$ and $c = -1$. Then $a + 2c + f = 1 > 0$ but
$$
r(x, \ov y) =
\begin{pmatrix}\ov y_1 & \ov y_2 \end{pmatrix}
\begin{pmatrix}
3 & 0 \\ 0 & -2
\end{pmatrix}
\begin{pmatrix} x_1 \\ x_2 \end{pmatrix},
$$
is neither positive nor negative semidefinite, so the scalar curvature
does not dominate the Ricci curvature.
\end{exam}

\begin{exam}[$r \not\to H$]
Let $a = f = 2$ and $c = -1$. Then
$$
r(x, \ov y) =
\begin{pmatrix}\ov y_1 & \ov y_2 \end{pmatrix}
\begin{pmatrix}
1 & 0 \\ 0 & 1
\end{pmatrix}
\begin{pmatrix} x_1 \\ x_2 \end{pmatrix}
$$
is positive-definite, but at $x = (x_1, x_1)$ we have
$$
R(x, \ov x, x, \ov x) = -4 |x_1|^2
$$
so the Ricci curvature does not dominate the holomorphic sectional curvature.
Note that the Ricci form is a multiple of the metric, so the metric is
K\"ahler--Einstein at the origin.
Any statement like ``this metric is K\"ahler--Einstein so its holomorphic
sectional curvature has a sign'' thus needs more information about the metric
or the manifold to have a chance of being true.
\end{exam}

\begin{exam}[$H \not\to r$]
We have
$$
R(x, \ov x, x, \ov x)
\geq
a |x_1|^2 + 2c |x_1|^2 |x_2|^2 + f |x_2|^4
$$
and the right-hand side is the quadratic form defined by
$$
Q = \begin{pmatrix}
a & c \\ c & f
\end{pmatrix}
$$
This form is positive-definite if and only if
$$
0 < \tr Q = a + f
\qandq
0 < \det Q = af - c^2.
$$
Suppose $c = -f - \epsilon$ with $\epsilon > 0$.
Then
$$
af - c^2
= af - f^2 - 2f \epsilon - \epsilon^2
= f(a - f - 2\epsilon) - \epsilon^2.
$$
If $a > f$, this is positive for small enough $\epsilon$ (like $a = 10$, $f =
1$, $\epsilon = 1$).
But then the Ricci form is
$$
r(x, \ov y) =
\begin{pmatrix}\ov y_1 & \ov y_2 \end{pmatrix}
\begin{pmatrix}
a & 0 \\ 0 & -\epsilon
\end{pmatrix}
\begin{pmatrix} x_1 \\ x_2 \end{pmatrix}
$$
which is not positive-definite,
so the holomorphic sectional curvature does not dominate the Ricci
curvature.
\end{exam}

\begin{exam}[$r \not\to R$]
We have
$$
\det R = acf - c^3 = c(af - c^2) = c \det r.
$$
It is then enough to pick a metric with $r$ positive-definite but $c < 0$
(like $a = f = 2$, $c = -1$)
to get a curvature tensor whose Ricci tensor is positive but is not itself
Griffiths positive, because it will satisfy $R(z, \ov w, z, \ov w) < 0$.
\end{exam}

\begin{exam}[$H \not\to R$]
Let $a = f = 2$ and $c = -1$.
Then
$$
R(x, \ov x, x, \ov x)
\geq |x|^4 + (|x_1|^2 - |x_2|^2)^2 > 0
$$
so the holomorphic sectional curvature is positive, but
for $x = (x,0)$ and $y = (0,y)$ we have
$$
q(x \odot y)
= R(x, \ov x, y, \ov y)
= - (|x|^2 + |y|^2 + x \ov y + y \ov x)
\leq 0
$$
and the inequality can be strict, so the holomorphic sectional curvature does
not dominate the Griffiths curvature.
\end{exam}


\subsection*{Known compact examples}


\begin{exam}[$H > 0 \not\to r > 0$ or $R > 0$]
As noted earlier, a Hirzebruch surface carries a K\"ahler metric with positive
holomorphic sectional curvature. There are such surfaces that do not have ample
canonical bundle, so they do not carry any K\"ahler metric with positive Ricci
curvature.
Those surfaces are also not the projective space, so they do not carry a
Griffiths-positive metric.
\end{exam}

We do not know of an example that shows that $H < 0 \not\to r < 0$ \emph{for
the given metric}.
By Wu--Yau--Tosatti--Yang it is probably difficult to find one.


\bibliographystyle{plain}
\bibliography{main}


\end{document}
